### 2024-03-31 Draft4.tex
Draft4.tex

\documentclass{article}

\usepackage[utf8]{inputenc}
\usepackage[T2A]{fontenc}
\usepackage[russian]{babel}
\usepackage[tmargin=1in,bmargin=1in,lmargin=1.25in,rmargin=1.25in]{geometry}

\usepackage[hidelinks,unicode]{hyperref}
\usepackage[threshold=0]{csquotes}
\usepackage{indentfirst}
\usepackage{enumerate}
\usepackage{totcount}
\usepackage{titlesec}
\usepackage{etoolbox}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hhline}
\usepackage{array}

\input{Common}

\begin{document}

\begin{lemma*}
    Всякий голоморфный $f \in \Hom((\torus)^n, \torus)$ является мономом, то~есть:
    $$
        \exists \alpha \in \Z^n{:}\ \forall z{:}\ f(z) = z^\alpha.
    $$
\end{lemma*}

\begin{proof}

Рассмотрим вещественный тор $T_\varepsilon = \{ z \in (\torus)^n\ |\ |z_1| = \varepsilon \wedge \ldots \wedge |z_n| = \varepsilon \}$.

???

\end{proof}

\end{document}